\chapter{Materials and Methods}

\section{Dataset}
The PlantVillage dataset is used in our study. PlantVillage is a collection of 54309 annotated diseased and healthy leaf images spanning several plant species, including tomato \cite{hughesOpenAccessRepository2016}. The size of the subset of tomato leaf images is 18162. Several images of each tomato leaf were captured outside under natural light in various lighting conditions and orientations, against a black or grey background, using a standard digital camera \cite{hughesOpenAccessRepository2016, kiranPlantLeafDisease2023}. The images are captured in the RGB colour space and have dimensions 256x256. We have selected PlantVillage as our primary dataset because it is widely used in image processing research for plant disease detection \cite{liReviewComputerVision2020, jiangConvolutionalNeuralNetworks2020, jafarRevolutionizingAgricultureArtificial2024, g.IdentificationPlantLeaf2019, francisDiseaseDetectionClassification2019, govardhanDiagnosisTomatoPlant2019}. Using this dataset will allow future researchers to compare our proposed method to previous studies.

\section{Image processing pipeline}
Many studies in image-based plant disease detection typically have predetermined phases in their image processing pipeline. Namely, preprocessing, segmentation, and feature extraction \cite{jafarRevolutionizingAgricultureArtificial2024, kaurPlantsDiseaseIdentification2019, perez-sanzPlantPhenomicsOverview2017, alhiaryFastAccurateDetection2011, desaiEnhancingImageQuality2025}. We follow suit and include each of those phases in our image processing pipeline.

\subsection{Preprocessing Phase}
Preprocessing involves tasks such as noise removal, image enhancement, and colour space conversion. In image acquisition, noise from the imaging sensor, the environment, or another source can compromise the accuracy of downstream image processing and classification phases. For example, we choose a global thresholding segmentation method, which we discuss later, and these methods often perform poorly in the presence of noise or uneven illumination \cite{gonzalezDigitalImageProcessing2017}. Nevertheless, preprocessing steps such as histogram equalisation and median filtering effectively mitigate noise and illumination issues \cite{gonzalezDigitalImageProcessing2017, saiImageQualityEnhancement2023, thangaduraiComputerVisionimageEnhancement2014}. Thus, preprocessing is a crucial phase of image processing pipelines, which enhances the overall accuracy of segmentation and image classification  \cite{kiranPlantLeafDisease2023, parkImagePreProcessingMethod2021}. Therefore, we apply several preprocessing steps to enhance the overall effectiveness and accuracy of our segmentation and classification phases.

\subparagraph{}
We aim to compute texture features of the tomato leaf in each image using the grey-level co-occurrence matrix \cite{haralickTexturalFeaturesImage1973}, which requires a grey-level image to calculate. Therefore, the first preprocessing step we perform is the conversion from the leaf image's original RGB colour space to the grayscale colour space. 

\subparagraph{}
Once we obtain the grayscale image of the tomato leaf, we apply histogram equalisation. This preprocessing method is used in the literature to add more contrast between the background and foreground of an image \cite{saiImageQualityEnhancement2023}. Image contrast enhancement is achieved by uniformly distributing the grey-level intensity values of the image \cite{gonzalezDigitalImageProcessing2017}, which makes histogram equalisation particularly useful for dealing with various illumination conditions \cite{kiranPlantLeafDisease2023}. 

\subparagraph{}
Lastly, we apply median filtering to remove common salt-and-pepper noise from the image. Median filtering is an effective image enhancement preprocessing step that has been successfully used in plant leaf disease detection \cite{hamudaSurveyImageProcessing2016a, desaiEnhancingImageQuality2025}. Noise manifests as sharp transitions in image intensity seen in the grayscale image histogram \cite{gonzalezDigitalImageProcessing2017}, therefore noise can negatively affect the accuracy of global threshold segmentation methods, which we use in this study; therefore, to  enhance segmentation accuracy and robustness, we apply median filtering to denoise the image.

\subsection{Segmentation Phase}
We use the Otsu method \cite{otsuThresholdSelectionMethod1979} for segmentation. The global thresholding Otsu method is used for segmentation. Global thresholding methods partition images into regions based on their intensity values using the image's grey-level histogram. For simple background removal, we select a suitable threshold T. If the intensity of the image at a point (x, y) exceeds T, then that part of the image is considered part of the foreground; otherwise, it is considered part of the background. If applied as a spatial-intensity transform, a binary mask is obtained where the intensity value of the foreground is set to 1, and the background is set to 0.  Global thresholding performs poorly when there is low contrast between the background and foreground, large amounts of noise, and when the background intensity varies significantly across the image \cite{rogowskaChapter5Overview2009}. More accurate segmentation can be achieved using adaptive thresholding \cite{jardimImageThresholdingApproaches2023}. Another option is learning-based segmentation, but it requires substantial amounts of training data to achieve reliable segmentation \cite{samalIntelligentImageAnalysis2021}. However, global thresholding is a simple, easy to implement, computationally efficient \cite{gonzalezDigitalImageProcessing2017, jardimImageThresholdingApproaches2023}. For this reason, we make the trade-off of poorer segmentation accuracy and robustness for computational efficiency. Moreover, global thresholding methods are applied in many studies in the literature for image-based plant phenotyping applications \cite{liReviewComputerVision2020}.

\subparagraph{}
The Otsu method automatically selects a threshold value that maximises the interclass variance of the intensity levels between the background and foreground in the image's grey-level histogram. This method is widely used in the literature \cite{liReviewComputerVision2020, najjarFlowerImageSegmentation2012}, and can be utilised for background removal in plant image processing pipelines \cite{perez-sanzPlantPhenomicsOverview2017}. However, the Otsu method is sensitive to noise and lacks robustness to uneven illumination \cite{liReviewComputerVision2020}. Despite the method's shortcomings, we have selected it because Otsu segmentation yields good segmentation results, is easy to implement, and is highly efficient \cite{najjarFlowerImageSegmentation2012, liReviewComputerVision2020}. 

\subparagraph{}
Concretely, for each leaf image, our segmentation phase aims to segment the leaf from its background. During feature extraction, we wish to obtain textural features of the leaf using information represented in the leaf's grayscale image. Therefore, background removal is crucial in ensuring that feature extraction considers the textural information of the leaf, rather than the background. The aforementioned Otsu segmentation method is applied to the grayscale leaf image obtained from the preprocessing phase. The result of the segmentation is a binary mask, and background removal is achieved by applying the binary mask to the preprocessed grayscale leaf image. 

\subsection{Feature Extraction Phase}
After segmentation, we have the isolated leaf image in the grey-scale colour space, and proceed to feature extraction. We are only extracting textural features from metrics calculated using a grey-level co-occurrence matrix \cite{haralickTexturalFeaturesImage1973}. Textural features are empirically some of the best predictors of leaf disease detection \cite{kimClassificationGrapefruitPeel, arivazhaganDetectionUnhealthyRegion2013}, since the occurrence of a plant disease induces specific changes in the texture of the leaf \cite{mokhtarSVMBasedDetectionTomato2015}. Thus, we choose textural features because we can construct feature vectors from these texture metrics to train a classifier to detect the presence or absence of a tomato leaf disease \cite{ahmedLeafImagebasedPlant2021, mokhtarSVMBasedDetectionTomato2015, arivazhaganDetectionUnhealthyRegion2013}.

\subparagraph{}
Gray-level co-occurrence matrices (GLCM) \cite{haralickTexturalFeaturesImage1973} measure the statistical spatial distribution of gray-level tones, and several parameters can be calculated from these matrices to investigate the textural properties of an object in an image \cite{argentiFastAlgorithmsTexture1990}. GLCM derived textural features are widely used in the literature to detect plant leaf diseases \cite{arivazhaganDetectionUnhealthyRegion2013, mokhtarSVMBasedDetectionTomato2015}.

\subparagraph{}
Following the approach in \cite{mokhtarSVMBasedDetectionTomato2015}, we compute a GLCM from the grayscale leaf image produced by the segmentation phase. Thereafter, we extract the following parameters from the GLCM: energy, contrast, sum of squares, correlation, entropy, sum entropy, cluster shade, cluster prominence, and homogeneity metrics \cite{arivazhaganDetectionUnhealthyRegion2013, haralickTexturalFeaturesImage1973, mokhtarSVMBasedDetectionTomato2015}. Together, all these calculated GLCM parameters form a feature vector, which we will use in our classification mechanism, as discussed in the next section.

\section{Classification  Mechanism}
For our classification mechanism, we use a support vector machine (SVM) \cite{vapnikNatureStatisticalLearning2000} to determine if a tomato leaf is diseased or healthy. SVM is a supervised machine learning method that has been shown to be an effective classifier of leaf diseases \cite{mokhtarSVMBasedDetectionTomato2015, kiranPlantLeafDisease2023, harakannanavarPlantLeafDisease2022, govardhanDiagnosisTomatoPlant2019}. SVM classifies several different classes by finding an optimal hyperplane that linearly separates them. SVMs can learn non-simple datasets and generalize well \cite{kiranPlantLeafDisease2023}.

\subparagraph{}
To obtain the training dataset, we apply our aforementioned image processing steps to obtain texture feature vectors. The feature vector is labeled as healthy if it is generated from a healthy PlantVillage tomato leaf image; otherwise, it is labeled as diseased.

\subparagraph{}
With this training dataset, we train our Support Vector Machine using the radial basis function kernel \cite{boolchandaniExploringEfficientKernel}. 
To avoid overfitting and bias from the dataset, we perform 10-fold cross-validation \cite{AdvancedDataMining2008}. In this training methodology, the data set is divided equally or nearly equally into 10 subsets. Then, cross-validation is performed 10 times, where one set is selected as the test set, and the union of the others forms the training set. The cross-validation procedure is repeated 10 times.

\section{Analysis}
We aim to measure the reliability of our binary SVM tomato classifier by evaluating its F-measure, accuracy, precision, recall, actual positive rate, and true negative rate \cite{AdvancedDataMining2008}. These metrics are commonly used to measure the performance of binary classifiers and predictive models \cite{lopez-garciaAutomaticDetectionSkin2010}; therefore, their use allows comparison of our system with other studies. Additionally, we measure the average classification time of the system, which is the time it takes for classification to occur from the start of preprocessing to the end of classification.

\paragraph{Precision}
Suppose there are $N$ samples, and a classifier $C$ predicts that a sample $x$ is positive and it is positive, then this prediction is called a true positive. If the classifier predicts $x$ is positive when it is not, then this prediction is called a false positive. We denote TP as the total number of true positives, and FP as the total number of false positives. Then precision is the ratio of all true positives over all positives observed. Symbolically,

\begin{equation}
	\text{Precision} = \frac{TP}{TP + FP}.
\end{equation}

\paragraph{Accuracy}
Using the terminology defined above accuracy is the ratio of all correct predictions over the total number of samples. Note that analoglosly the TN is a prediction for which a classifier correctly predicts a sample is negative. And FN is the converse that is if the sample is predicted as negative when it is not. Then symbolically accuracy is defined as: 

\begin{equation}
	\text{Accuracy} = \frac{TP + TN}{TP + TN + FP + FN}.
\end{equation}

\paragraph{Recall}
Also known as true positive rate or a classifiers ablity to correctly identify true positive samples is defined as:

\begin{equation}
	\text{Recall} = \frac{TP}{TP +  FN}.
\end{equation}

\paragraph{F1-score}
F1-score is the harmonic mean or weighted avergae of precision and recall. F1-score takes false positive and false negative predictions into account. Symbolically:

\begin{equation}
	\text{F1} = 2* \frac{Precision  * Recall}{Precision +  Recall}.
\end{equation}