\chapter{Introduction}

\paragraph{}
The rapid growth of the human population has created a strain on international food security. Consequently, crop production must increase by 70\% by 2050 to meet this demand \cite{carvajal-yepesGlobalSurveillanceSystem2019}. Over the past decade, methods in image-processing-based plant phenotyping have been employed to investigate plant gene-environment interactions in greenhouse environments to develop cultivars that exhibit optimal yield and resistance to abiotic and biotic stressors. Despite these advances, current solutions are non-generalizable to a broad range of plant species and can be costly \cite{arnalbarbedoDigitalImageProcessing2013, liReviewImagingTechniques2014, minerviniImageAnalysisNew2015}. Thus, there is a critical need for affordable, low-cost plant phenotyping image processing tools in automated greenhouses. This study develops an open-source image-processing-based plant phenotyping component, which helps to detect various factors of plant health, stress and growth in order to support low-cost automated greenhouses. The low-cost imaging component developed utilises low-complexity and efficient image segmentation techniques suitable for resource-constrained hardware, leveraging open-source software and publicly available labelled plant image datasets. Additionally, our component is extensible and flexible, making it more accessible and generalisable to other plant environment contexts. Following this introduction, we review the literature on image processing in greenhouse environments, paying particular emphasis to plant phenotyping and its applications with image processing in automated, low-cost greenhouse environments. Thereafter, we discuss our methodology for developing the low-cost image processing tool, detailing our automated image processing and analysis pipelines. Lastly, we discuss the effectiveness and cost-effectiveness of the developed component using publically available plant image datasets. 