% The problem of plant diseases
\paragraph{The problem of plant diseases}
Our lives would not be possible without agriculture. We use goods like medicine, pesticides, and rubber everyday, all of which are made from plants. Plants are ubiquitous in our lives, and they fulfill some of our essential needs. Most importantly, plants feed most people globally \cite{mccouchFeedingFuture2013}. But many of us still suffer from hunger. In 2024, it was estimated that between 638 and 720 million people did not have enough to eat \cite{StateFoodSecurity2025}. Plant dieases make this situation even worse by robbing us of food \cite{carvajal-yepesGlobalSurveillanceSystem2019, heHighThroughputRoboticPhenotyping2025, nazarovInfectiousPlantDiseases2020, hughesOpenAccessRepository2016}. Therefore, they are a direct threat to our livlihoods.
\subparagraph{}
Plant diseases make it difficult for crops to survive and stay healthy \cite{agriosINTRODUCTION2005}. They interfere with a plant's cells causing them to malfunction or die \cite{huberRelationshipNutritionPlant2012}. And each cell helps the plant carry out essential life-sustaining functions: conducting photosynthesis, and absorbing water and nutrients. So, when these cells are damaged, the plant's ability to obtain and use nutrients is limited. As a result, plant diseases stunt the growth, and decrease the yield and nutritional value of crops \cite{savaryCropLossesDue2012, vyskaTradeoffDiseaseResistance2016, chinPlantDiseaseDetection2023, kitpoEarlyRiceDisease2018, martinelliAdvancedMethodsPlant2015}. In some cases, the damage is so severe that farmers lose up to 100\% of their crops \cite{nazarovInfectiousPlantDiseases2020, dongPlantPADPlatformLargescale2024}. However, the losses caused by plant diseases goes further than just the yield and quality of crops; they extend to rural communities, consumers, the environment, and governments \cite{savaryCropLossesDue2012}. Therefore, plant diseases threaten the health of crops, and this has disastrous far-reaching effects on our livlihoods. Clearly, we need solutions to deal with them.
s
\paragraph{Some solutions to plant disease}
Luckily, several options exist. For example, plant breeding is an environmentally sustainable and cost-effective solution, which helps to produce disease-resistant varieties of plants \cite{anandBreedingTechniquesDispense2023, ortonIntroduction2020, heHighThroughputRoboticPhenotyping2025, furbankPhenomicsTechnologiesRelieve2011}. Alternatively, we also resort to precision agriculture, which uses data from crops and their environment to make optimal decesions about agricultural actions \cite{PrecisionAgricultureFood2025, PlantDiseaseDetection, cisternasSystematicLiteratureReview2020, yandunnarvaezSurveyRangingImaging2017}. For instance, precision agriculture aids farmers in deciding on the best way to apply pesticides during a pathogen epidemic \cite{bockPlantDiseaseSeverity2010} by helping determine when to apply pesticides, how much pesticide to apply, and where to apply it; as a result, such precise pesticide application will reduce costs, save resources, and benefit the environment and human health \cite{pathakCurrentStatusPesticide2022, reiser3DImagingSonar2017, khalidRealTimePlantHealth2023, qiaoEditorialAISensors2022, dechantAutomatedIdentificationNorthern2017, HelpingFarmersFace2025}. So, plant breeding and precision agriculture help deal with plant diseases; however, these methods depend on effective plant phenotyping \cite{fahlgrenVersatilePhenotypingSystem2015, furbankPhenomicsTechnologiesRelieve2011a, gehanPlantCVV2Image2017, zhangBOCNNBiLSTMDeepLearning2024, zhuDeepLearningbasedText2025a, anandBreedingTechniquesDispense2023}.

\paragraph{The relationship between plant diseases and plant phenotyping}
Plant phenotyping is the evaulation of observable or measurable plant phenotypes \cite{fioraniFutureScenariosPlant2013, johannsenGenotypeConceptionHeredity1911a}, which are proxy measurements or traits indicating a plant's stress, health or growth in relation to its environment \cite{fioraniFutureScenariosPlant2013, liReviewImagingTechniques2014}. Moreover, phenotypes represent traits of an individual plant in terms of all of its chemichal, physilogical, morphological, and behavioral aspects \cite{guoPhenotypingPlants2014, mohammedReviewPlantPhenotyping2025, liReviewImagingTechniques2014}. So, plant phenotyping provides quantative data that represents various aspects of a plant. When a plant is diseased its form, function and integrity are affected \cite{agriosINTRODUCTION2005}, therefore reliable measurments of all aspects of plant are required in plant disease solutions. Consequently, plant phenotyping is an effective quantative method for plant breeding and precision agriculture \cite{gehanPlantCVV2Image2017, wenThermalImagingDigital2023, dongPlantPADPlatformLargescale2024, mutkaImagebasedPhenotypingPlant2015a, anandBreedingTechniquesDispense2023}.

\paragraph{Plant phenotyping and plant breeding}
The goal of plant breeding is to develop plants with desirable traits like disease resistance. In this regard, the most important tasks for breeders are the selection of suitable individual plants and the developement of genetic material \cite{anandBreedingTechniquesDispense2023}. Plant phenotyping assists in these tasks. It does so by providing breeders data about a plant's phenotype-environment interactions, which allows them to identify and select suitable parameters for the development of improved crops \cite{xiaoAdvancedHighthroughputPlant2022}. And this method has been used successfully in several breeding programmes for disease resistance. To name a few, \cite{visioniGenomeWideAssociation2018} conducted a study for Barley Rust Stripe resistance, \cite{zhuIdentificationQuantitativeTrait2003} for Oat Crown Rust resistance, \cite{pumphreyValidatingFhb1QTL2007} for Wheat Head Blight resistance, \cite{mcclureGenomeWideAssociationStudy2018} for Apple Scab resistance, and \cite{yangGbvdr6GeneEncoding2018} for Cotton Verticillium Wilt resistance. Thus, plant phenotyping effectively assists plant breeding programmes.

\paragraph{Plant phenotyping and precision agriculture}
Precision agriculture provides data-driven decision making for optimal agricultural management, and plant phenotyping can enable this \cite{zhuDeepLearningbasedText2025, qiaoEditorialAISensors2022, reiser3DImagingSonar2017, dechantAutomatedIdentificationNorthern2017, bergerPrecisionAgricultureEnabler2022}. Notably, there are many pesticide, disease detectionl and warning systems that rely on phenotyping \cite{khalidRealTimePlantHealth2023, kitpoEarlyRiceDisease2018}. For example, \cite{caffiEvaluationWarningSystem2010} discusses a warning system for controlling infections of Grapevine Downy Mildew, and \cite{bergerZoneSelectivePlantProtection2022} presents an early disease detection system for selective pesticide application which successfully helped managed diseases in apple, olive and almond farms. Evidently, precision agriculture with phenotyping provides effective plant disease management. But traditional phenotyping has problems.

\paragraph{Tradtional plant phenotyping solutions, and their challenges}
Human naked eye observation is the main traditional phenotyping method used in plant disease detection. This method needs continous effort from human raters which can be expensive. And in poorer countries, raters are far and few between \cite{rameshPlantDiseaseDetection2018, basavaiahTomatoLeafDisease2020}, so farmers will need to expend more time, energy and resources to obtain raters \cite{alhiaryFastAccurateDetection2011, DetectionClassificationLeaf2025, gillComprehensiveReviewHigh2022}. Thus the consultation of human experts is expensive, time-consuming and not accessible to everyone. Additionally, naked eye observation has practical limitations. For example, when plant samples are needed, visual assessment is destructive, and this prevents the observation of plant traits over time \cite{gehanPlantCVV2Image2017, gillComprehensiveReviewHigh2022, hangClassificationPlantLeaf2019}. Moreover, visual assessment is error prone when a disease does not induce any visible symptoms in a plant, thus the detection of early disease onset is difficult for human raters. However, alternative labratory methods for early disease detection exist \cite{arnalbarbedoDigitalImageProcessing2013}, but they can also be computationally expensive and time consuming just like human visual assessment \cite{basavaiahTomatoLeafDisease2020, rameshPlantDiseaseDetection2018, bockPlantDiseaseSeverity2010, PlantDiseaseDetection}. Fortunatley, image sensing technologies combined with machine learning and image processing techniques can overcome many problems that traditional phenotyping methods suffer from.

\paragraph{Image sensing technlogies an alternative to traditional phenotyping methods}
Image sensing technologies overcome some of the deficiencies of traditional plant phenotyping methods. For example, traditional methods can be destructive, which complicates the measurement of plant traits over time. And measuring phenotypes over time is important because plant disease symptoms can be latent or appear suddenly, and then progress rapidly. Thus, non-destructive phenotyping methods are beneficial for early disease detection and treatment. Evidently, traditional methods are lacking in this regard, but image sensing technologies are not, since they can precisely and non-destrutively sample large populations. Moreover, this can be done in a very short period of time. Therefore, image sensing technology is more economical than time-consuming human labour. 

\subparagraph{}
Sensors also capture plant tissue–light interactions which allows for  the measurement and quantification of disease symptoms that are not visible to the eye. Specifically, plant tissues absorb, reflect, emit, transmit, and fluoresce different wavelengths of light. Light is captured by a sensor, and this light can have different wavelengths each of which may correspond to distinct biochemical or physiological aspects of a plant that are not visible to humans. Therefore, sensing technlogies increase the possible types of traits we can observe, which allows for earlier, more accurate, and comprehensive disease symptom detection. 

\subparagraph{}
Numerous studies have demonstrated the effectiveness of sensing technlogy in plant disease detection and management.  \cite{burlingQuantumYieldNonregulated2010} used chlorophyll fluorescence imaging to detect leaf rust and assess its severity in wheat.  \cite{chaerleThermalChlorophyllFluorescenceImaging2004} used thermography and chlorophyll fluorescence imaging to non-destructively visualize \emph{Cercospora beticola} infections. \cite{yanyangEarlyDetectionRice2012} applies near-infrared hyper-spectral image  to detect blast rice in Nipponbare seedlings. \cite{xieDetectionEarlyBlight2015} used hyperspectral imaging to detect early blight and late blight diseases in tomato leaves. \cite{yangTeaDiseasesDetection2019} uses thermal image sensing to detect tea diseases. Clearly, image sensors have proven effective in supporting plant disease detection and management. Furthermore, in recent years, a wide range of machine learning and image processing techniques have been used with these sensors.

\paragraph{Image processing pipelines}
% This paragraph describes image processing pipelines in the phases of preprocessing, segmentation, and classification/detection. The aim here is to collect the general methods applied and discuss various applications of them in tomatoe leaf disease detection 

\clearpage

%\paragraph{}
%In this section, the literature related to image-processing in tomato plant pathogen detection is reviewed. First a high-level overview of plant phenotyping is given, since plant phenotyping is foundational to image-processing methods in plant disease detection. Image-based phenotyping methods are then discussed, highlighting the advantages these methods have over traditional ones. Next, a classification of these solutions into three types is presented: traditional, machine learning, and deep learning solutions. And this classification is used to evaluate current image-processing solutions for plant pathogen detection across different crops, with emphasis on tomato plants. Finally, the main challenges and limitations of these methods are identified and possible solutions are presented.

%\paragraph{}
%This literature review provides a foundational understanding of image processing in plant phenotyping and directions for developing or investigating solutions for low-cost image-processing phenotyping tools in automated greenhouses. The focus is on the existing image processing solutions applied in low-cost automated greenhouse environments. We begin by examining the fundamental concepts of plant phenotyping and the various traits that are amenable to image processing for predicting plant health, growth, and stress. Specifically, we note that leaf phenotypic traits are a robust proxy indicator of physiological and metabolic factors affecting plant health \cite{zhangHighthroughputPhenotypingPlant2023}.

%\paragraph{}
%Thereafter, aspects of image processing in plant phenotyping are discussed. Notably, image processing and acquisition, as well as various image sensor technologies and segmentation methods, are reviewed in the context of plant phenotyping. Hyperspectral and RGB imaging are the most common acquisition methods, and segmentation techniques critically influence the overall accuracy, precision, and quality of phenotypical trait analysis \cite{choudhurySegmentationTechniquesChallenges2020, muller-linowPlantScreenMobile2019}.

%Lastly, the status, challenges, and potential innovations of low-cost automated image processing in greenhouse environments are discussed. In this vein, the literature indicates a lack of affordable and generalizable image processing tools that can analyse and process large amounts of image data on inexpensive, resource-constrained hardware and sensors \cite{gomesComputationalSustainabilityComputational2009, minerviniImageAnalysisNew2015}.